---
title: "Central Limit Theorem to Exponential Distribution Comparison"
author: "Sebastian Cruz"
date: "December 18, 2015"
output: html_document
---

## Overview

## Simulations
Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes

## Sample Mean vs Theoretical Mean
Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers. 
 
## Sample Variance vs Theoretical Variance
Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances. 
 
## Distribution
Via figures and text, explain how one can tell the distribution is approximately normal.
focus on the difference between the distribution of a large collection of random exponentials and the distribution of a large collection of averages of 40 exponentials. 

```{r}
summary(cars)
```

You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
